---
title: "Allergy Dataset"
author: "Silvia Garces"
date: "`r Sys.Date()`"
output: html_document
---

# ONLY METABOLOMICS + PROTEOMICS 46T
Now we will try to see if there is any division in T0, that means, check if the methods can separate the phenotypes of the patients at baseline.
This dataset has the samples in rows while variables are in columns

###1. SNF
Samples in rows and variables in columns
```{r}
library(SNFtool)
Sys.time()
#Calculate distances within the datasets
dist_metabo_T0 = (dist2(as.matrix(omics_T0$Metabo_T0), as.matrix(omics_T0$Metabo_T0)))^(1/2)
dist_prot48_T0 = (dist2(as.matrix(omics_T0$Prot48T_T0), as.matrix(omics_T0$Prot48T_T0)))^(1/2)

K=20
sigma=0.5

W1 = affinityMatrix(dist_metabo_T0, K, sigma)
W2 = affinityMatrix(dist_prot48_T0, K, sigma)

W_joined <- SNF(list(W1,W2), K, 20)

estimateNumberOfClustersGivenGraph(W_joined, 2:5)
Sys.time()
```
Select optimal number of clusters
```{r}
C=2 #number of clusters
group=spectralClustering(W_joined,C)

displayClusters(W_joined,group)

#Check with true labels:
M_label=cbind(group, as.numeric(as.factor(samples_df_T0$treatment)))
colnames(M_label)=c("spectralClustering","TrueLabel")
rownames(M_label)=samples_df_T0$sample
M_label_colors=cbind("spectralClustering"=getColorsForGroups(M_label[,"spectralClustering"],
colors=c("blue","red")),"TrueLabel"=getColorsForGroups(M_label[,"TrueLabel"], colors = c("green","violet")))

#Plot heatmap
layout(matrix(c(1,2), ncol=2, byrow=TRUE))
par(mar=c(5, 4, 4, 2))
# Plot the heatmap
displayClustersWithHeatmap(W_joined, group, M_label_colors)

par(mar=c(0,0,0,0))
plot(1, type = "n", axes=F, xlab="", ylab="") # Create empty plot
legend(x="top", legend=unique(group), fill=unique(M_label_colors[,"spectralClustering"]), title = "SNF clusters", bg="transparent", horiz=TRUE)
legend(x="center", legend=unique(samples_df_T0$treatment), fill= c("green","violet"), title = "Treatment", bg="transparent", horiz=TRUE)
```
###2. NEMO:
Data needs to be transposed as samples need to be in columns
```{r}
library(NEMO)
Sys.time()
#transpose matrices
data_transposed_2_omics <- list(t(omics_T0$Metabo_T0),t(omics_T0$Prot48T_T0))

affinity_graph_T0_2 <- nemo.affinity.graph(data_transposed_2_omics, k= 20)
num_clusters_T0_2 <- nemo.num.clusters(affinity_graph_T0_2)
clustering_nemo_T0_2 <- nemo.clustering(data_transposed_2_omics, num.clusters = num_clusters_T0_2, num.neighbors = 20)

NEMO_label=cbind(clustering_nemo_T0_2, as.numeric(as.factor(samples_df_T0$treatment)))
colnames(NEMO_label)=c("nemoClustering","TrueLabel")
rownames(NEMO_label)=samples_df_T0$sample

NEMO_label_colors=cbind("nemoClustering"=getColorsForGroups(NEMO_label[,"nemoClustering"],
colors=c("blue","red")),"TrueLabel"=getColorsForGroups(NEMO_label[,"TrueLabel"], colors = c("green","violet")))

displayClustersWithHeatmap(affinity_graph_T0_2, clustering_nemo_T0_2, NEMO_label_colors)

plot.new()
legend(x="top", legend=unique(clustering_nemo_T0_2), fill=unique(NEMO_label_colors[,"nemoClustering"]), title = "NEMO clusters", bg="transparent", horiz=TRUE, cex=0.8)
legend(x="center", legend=unique(samples_df_T0$treatment), fill= c("green", "violet"), title = "Treatment", bg="transparent", horiz=TRUE, cex=0.8)
Sys.time()
```
###3. MCIA:
Use transposed data
```{r}
library(omicade4)
Sys.time()
mcoin_2 <- mcia(data_transposed_2_omics)
plot.mcia(mcoin_2,axes = 1:2 , sample.lab = FALSE, df.color = 1:2)
plotVar(mcoin_2, var=NA, bg.var.col=1:2, var.lab=TRUE)
Sys.time()
```

Make plots using 2 latent variables
```{r}
color <- c("lightblue3","maroon")
subset_mcia <- mcoin_2$mcoa$SynVar
par(mar=c(5, 4, 4, 8), xpd=TRUE)
plot(subset_mcia[, 1], subset_mcia[,2], xlab= "Variable 1", ylab= "Variable 2", col=color[as.numeric(as.factor(samples_df_T0$treatment))], pch=20, main = "Plot of samples in T0")
legend("right", legend=unique(samples_df_T0$treatment), col=color,pch=20, title="Treatment", inset=c(-0.3, 0))
```

###4. moCluster
Use transposed data
```{r}
library(mogsa)
Sys.time()
moa_2 <- mbpca(data_transposed_2_omics, ncomp=10, k="all", method = "globalScore", option = "lambda1",center=FALSE, scale= FALSE, moa= TRUE, svd.solver = "fast", maxiter = 1000)
plot(moa, value="eig", type=2)
#Check number of components/latent variables
r <- bootMbpca(moa_2, mc.cores = 1, B=20, replace = FALSE, resample = "sample")

```
Keep appropriate number of components = 2
```{r}
moa_2var_2 <- mbpca(data_transposed_2_omics, ncomp=2, k=0.25, method = "globalScore", option = "lambda1",center=FALSE, scale= FALSE, moa= TRUE, svd.solver = "fast", maxiter = 1000)
scr <- moaScore(moa_2)
scrs <-moaScore(moa_2var_2)

#do correlation
diag(cor(scr[, 1:2], scrs))

#Do hierarchical cluster and plot
hcl <- hclust(dist(scrs))
cls <- cutree(hcl, k=2) #Pre-select 2 clusters

Sys.time()
#Plot the results obtained with 2 components
clsColor <- as.factor(cls)
levels(clsColor) <- c("red","blue")
clsColor<- as.character((clsColor))
```
Plot results:
```{r}
#Plot heatmap and add legend:
heatmap(t(scrs[hcl$order, ]), ColSideColors = clsColor[hcl$order], Rowv = NA, Colv=NA)
legend(x= "right", legend = c("Cluster 1","Cluster 2"), col = c("red","blue"), pch= 15)

#Plot in component space
symbols <- c(16, 3)
par(mar=c(5, 4, 4, 8), xpd=TRUE)
plot(scrs[, 1:2], col=clsColor, pch=symbols[as.numeric(as.factor(samples_df_T0$treatment))], main = "PC1 and PC2")
legend("topright", legend=c("Cluster 1","Cluster 2"), col=unique(clsColor), pch=15, title="Clusters", inset=c(-0.25, 0))
legend("bottomright", legend=unique(samples_df_T0$treatment), col=c("lightslategrey"), pch=symbols, title="Treatment", inset=c(-0.3, 0))
```
### 5.PINSPlus:
Data can be used as original
```{r}
library(PINSPlus)
Sys.time()
omics_t0_2 <- list(omics_T0$Metabo_T0, omics_T0$Prot48T_T0)
pins_cluster_t0_2 <- SubtypingOmicsData(dataList =omics_t0_2 , clusteringMethod = "kmeans", kMax=3)

cluster1 <- pins_cluster_t0_2$cluster1
cluster2 <- pins_cluster_t0_2$cluster2

# Plot the clusters for Stage I using a scatter plot
plot(cluster1, pch = 16, col = "blue", main = "Cluster Plot (Stage I)", xlab = "Sample", ylab = "Cluster")
# Plot the clusters for Stage II using a scatter plot
plot(cluster2, pch = 16, col = "red", main = "Cluster Plot (Stage II)", xlab = "Sample", ylab = "Cluster")
Sys.time()

```
Plot PCA for first 2 components to check distribution of samples:
```{r}
# Subset the first two columns of the 'x' matrix (which are the PCA)
subset_x <- pins_cluster_t0_2[["dataTypeResult"]][[1]][["pca"]][["x"]]
cluster_colors <- c("coral", "cadetblue3")
symbols <- c(16, 3)

# Plot PC1 vs PC2
par(mar=c(5, 4, 4, 8), xpd=TRUE)
plot(subset_x[, 1], subset_x[, 2], 
     xlab="PC1", ylab="PC2", 
     main="Plot of PC1 vs PC2 in T0", 
     col=cluster_colors[cluster1],
     pch=symbols[as.numeric(as.factor(samples_df_T0$treatment))])
legend("topright", legend=c("Cluster 1","Cluster 2"), col=cluster_colors, pch=15, title="Clusters", inset=c(-0.25, 0))
legend("bottomright", legend=unique(samples_df_T0$treatment), col="lightslategrey", pch=symbols, title="Subtypes", inset=c(-0.3, 0))
```
###6.intNMF
Make values positive if there is any negative value in the matrices
```{r}
library(IntNMF)
Sys.time()
if (!all(omics_T0$Metabo_T0>=0)){
  data1_t0 <- pmax(omics_T0$Metabo_T0 + abs(min(omics_T0$Metabo_T0)), .Machine$double.eps)
} else {
  data1_t0 <- omics_T0$Metabo_T0
}

if (!all(omics_T0$Prot48T_T0>=0)){
  data2_t0 <- pmax(omics_T0$Prot48T_T0 + abs(min(omics_T0$Prot48T_T0)), .Machine$double.eps)
} else {
  data2_t0 <- omics_T0$Prot48T_T0
}

dat_t0 <- list(as.matrix(data1_t0), as.matrix(data2_t0))

#Get optimum number of clusters: nmf.opt.k
opt_k<- nmf.opt.k(dat=dat_t0, n.runs = 5, n.fold = 5, k.range = 2:7, result = TRUE, 
                  make.plot = TRUE, progress = TRUE)
```
Number of clusters: 2
```{r}
#We check and the optimal is 3
fit <- nmf.mnnals(dat=dat_t0, k=2, maxiter = 200, st.count = 20, n.ini = 15, ini.nndsvd = TRUE, seed = TRUE)

#Make silhouette scores
SilhouettePlot(fit, cluster.col = NULL)
ConsensusMatPlot(fit, colLab = TRUE, rowLab = TRUE)
Sys.time()
```
Plot the clusters obtained in different factors:
```{r}
clusters_int <- fit$clusters
subsets_df <- fit$W #get the basis matrix for the samples (nº of vectors = k)
clus_colors <- c("darkorange1", "deepskyblue")
symbols <- c(16, 3)

par(mar=c(5, 4, 4, 8), xpd=TRUE)
plot(subsets_df[, 1], subsets_df[, 2], 
     xlab="Component 1", ylab="Component 2", 
     main="Plot of Samples in space in T0", 
     col=clus_colors[clusters_int],
     pch= symbols[as.numeric(as.factor(samples_df_T0$treatment))])
legend("topright", legend=c("Cluster 1","Cluster 2"), col=clus_colors, pch=15, title="Clusters", inset=c(-0.25, 0))
legend("bottomright", legend=unique(samples_df_T0$treatment), col="lightslategrey", pch=symbols, title="Treatment", inset=c(-0.3, 0))
```

###7.MixOmics:
```{r}
library(mixOmics)
Sys.time()
names(omics_t0_2) <- c("metabo", "prot46") #list made in PINSplus method

#sPLS basic model
spls_2omics <- spls(X=omics_t0_2$metabo,Y=omics_t0_2$prot46, ncomp = 5, mode = 'regression')
plotIndiv(spls_2omics, group = samples_df_T0$treatment, ind.names= FALSE, legend = TRUE, legend.title = "Treatments")
```
Tune the model:
```{r}
# repeated CV tuning of component count
perf.spls.2omics <- perf(spls_2omics, validation = 'Mfold',
                         folds = 10, nrepeat = 5) 

plot(perf.spls.2omics, criterion = 'Q2.total')
```
Select number of variables:

```{r}
# set range of test values for number of variables to use from X dataframe
list.keepX <- c(seq(10, 30, 5))
# set range of test values for number of variables to use from Y dataframe
list.keepY <- c(seq(10, 30, 5)) 

#tunning parameters
tune.spls.2omics <- tune.spls(X=omics_t0_2$metabo, Y=omics_t0_2$prot46, ncomp = 2,
                              test.keepX = list.keepX,
                              test.keepY = list.keepY,
                              nrepeat = 1, folds = 10, # use 10 folds
                              mode = 'regression', measure = 'cor') 
plot(tune.spls.2omics)         # use the correlation measure for tuning
```
Get optimal:
```{r}
# extract optimal number of variables for X dataframe
optimal.keepX <- tune.spls.2omics$choice.keepX 

# extract optimal number of variables for Y datafram
optimal.keepY <- tune.spls.2omics$choice.keepY

optimal.ncomp <-  length(optimal.keepX)
```

Make final model:
```{r}
# use all tuned values from above
final.spls.2omics <- spls(X=omics_t0_2$metabo, Y=omics_t0_2$prot46, ncomp = optimal.ncomp, 
                    keepX = optimal.keepX,
                    keepY = optimal.keepY,
                    mode = "regression")
Sys.time()
```

Make plots:
```{r}
##X-variate
plotIndiv(final.spls.2omics, ind.names = FALSE, 
         rep.space = "X-variate", # plot in X-variate subspace
         group = samples_df_T0$treatment, # colour by time group
         col.per.group = color.mixo(1:2), 
         legend = TRUE, legend.title = 'Treatment')

##Y-variate
plotIndiv(final.spls.2omics, ind.names = FALSE,
         rep.space = "Y-variate", # plot in Y-variate subspace
         group = samples_df_T0$treatment, # colour by time group
         col.per.group = color.mixo(1:2), 
         legend = TRUE, legend.title = 'Treatment')

##XY-variate
plotIndiv(final.spls.2omics, ind.names = FALSE,
         rep.space = "XY-variate", # plot in Y-variate subspace
         group = samples_df_T0$treatment, # colour by time group
         col.per.group = color.mixo(1:2), 
         legend = TRUE, legend.title = 'Treatment')
```
Plot in 3D
```{r}
plotArrow(final.spls.2omics, ind.names = FALSE,
          group = samples_df_T0$treatment, # colour by time group
          col.per.group = color.mixo(1:2),
          legend.title = 'Treatmentp')
```
####Try DIABLO:
```{r}
#Specify Y which are the treatments
Sys.time()
Y<- samples_df_T0$treatment

#Do basic model
result.diablo.t0_2 <- block.plsda(omics_t0_2, Y) # run the method
plotIndiv(result.diablo.t0_2, legend = TRUE) # plot the samples
plotVar(result.diablo.t0_2) # plot the variables
```
Make another model
```{r}
design = matrix(0.1, ncol = length(omics_t0_2), nrow = length(omics_t0_2), 
                dimnames = list(names(omics_t0_2), names(omics_t0_2)))
diag(design) = 0 # set diagonal to 0s

# form basic DIABLO model
basic.diablo.model = block.splsda(X = omics_t0_2, Y = Y, ncomp = 5, design = design) 

# run component number tuning with repeated CV
perf.diablo = perf(basic.diablo.model, validation = 'Mfold', 
                   folds = 5, nrepeat = 10) 

plot(perf.diablo) # plot output of tuning


# set the optimal ncomp value
ncomp = perf.diablo$choice.ncomp$WeightedVote["Overall.BER", "centroids.dist"] 

# show the optimal choice for ncomp for each dist metric
#perf.diablo$choice.ncomp$WeightedVote 

# set grid of values for each component to test
test.keepX = list (metabo = c(seq(10,25,5)), 
                   prot48 = c(seq(10,25,5)))

# run the feature selection tuning
tune.omics_t0 = tune.block.splsda(X = omics_t0_2, Y = Y, ncomp = ncomp, 
                              test.keepX = test.keepX, design = design,
                              validation = 'Mfold', folds = 10, nrepeat = 1,
                              dist = "centroids.dist")
```
Best with 3 components:
```{r}
list.keepX = tune.omics_t0$choice.keepX # set the optimal values of features to retain
# set the optimised DIABLO model
final.diablo.model = block.splsda(X = omics_t0_2, Y = Y, ncomp = 3, keepX = list.keepX, design = design)
Sys.time()

#Make plots
plotDiablo(final.diablo.model, ncomp = 1)
plotIndiv(final.diablo.model, ind.names = FALSE, legend = TRUE, 
          title = 'DIABLO Sample Plots')
#X11()
#cimDiablo(final.diablo.model)
```
Plot AUC-ROC:
```{r}
auc.splsda_met = auroc(final.diablo.model, roc.block = "metabo", 
                  roc.comp = 2, print = FALSE)
auc.splsda_prot = auroc(final.diablo.model, roc.block = "prot46", 
                  roc.comp = 2, print = FALSE)
```

###8.MOFA
We need to use the transposed the data
```{r}
library(MOFA2)
Sys.time()
data_mofa_t0 = list(data_transposed_2_omics[[1]],data_transposed_2_omics[[2]])
names(data_mofa_t0)<- c("metabo", "prot48")
mofa_obj <- create_mofa(data_mofa_t0)
plot_data_overview(mofa_obj)

#Specify the model options
##Default
data_opts <- get_default_data_options(mofa_obj)

model_opts <- get_default_model_options(mofa_obj)
model_opts$num_factors <- 10

train_opts <- get_default_training_options(mofa_obj)
train_opts$seed <- 42

#Start training the model
mofa_obj <- prepare_mofa(mofa_obj, data_options = data_opts,
                            model_options = model_opts, training_options = train_opts)
outfile = file.path(getwd(),"model_10_6.hdf5") #save model
MOFA_object_trained <- run_mofa(mofa_obj, outfile, use_basilisk=TRUE)
Sys.time()
```
Make plots:
```{r}
plot_factor_cor(MOFA_object_trained)
##max_r2= max variance 
plot_variance_explained(MOFA_object_trained, max_r2 = 25)
##total variance
plot_variance_explained(MOFA_object_trained, plot_total = T)[[2]]
```
Factor 3 explains both of them
```{r}
plot_weights(MOFA_object_trained, view = "metabo", factor= 3, nfeatures = 10)
plot_weights(MOFA_object_trained, view = "prot48", factor= 3, nfeatures = 10)
```
Plot heatmaps
```{r}
plot_data_heatmap(MOFA_object_trained, view = "metabo", factor = 3, features = 15,
                  denoise = TRUE, cluster_rows= TRUE, cluster_cols= T, 
                  show_rownames= TRUE, show_colnames=TRUE  ,scale="row")
plot_data_heatmap(MOFA_object_trained, view = "prot48", factor = 3, features = 15,
                  denoise = TRUE, cluster_rows= TRUE, cluster_cols= T,
                  show_rownames= TRUE, show_colnames=TRUE  ,scale="row")
```
Now lets make clusters
```{r}
clusters_mofa_all <- cluster_samples(MOFA_object_trained, k=2, factor=c(2,3,6)) # we want 2 clusters
clusters_mofadf_all <- data.frame(cluster=clusters_mofa_all$cluster)

```

Check results- difficult to plot
```{r}
subset_fact <- MOFA_object_trained@expectations$Z[["group1"]]
colors_mofa <- c("darkorchid1","darkseagreen4")
symbol <- c(16,8)
par(mar=c(5, 4, 4, 8), xpd=TRUE)
plot(subset_fact[,2], subset_fact[,3], xlab="Factor 2", ylab= "Factor 3", main="Plot Factor 2 vs Factor 3 in T0", pch=symbol[as.numeric(as.factor(samples_df_T0$treatment))], col=colors_mofa[clusters_mofadf_all$cluster])
legend("topright", legend=c("Cluster 1","Cluster 2"), col=colors_mofa, pch=15, title="Clusters", inset=c(-0.25, 0))
legend("bottomright", legend=unique(samples_df_T0$treatment), col="lightslategrey", pch=symbol, title="Treatment", inset=c(-0.3, 0))
```
###9.iClusterPlus:

Samples need to be in rows so we do not need to modify the data
```{r}
library(iClusterPlus)
library(gplots)
library(latticeExtra)
Sys.time()
#try different k values and then compare the results
#Clusters= k+1
fit_c_1k <- iClusterPlus(dt1 = omics_t0_2$metabo, dt2 = omics_t0_2$prot46, type = c("gaussian","gaussian"), lambda = c(0.10,0.25), K=1, maxiter = 15)
fit_c_1k$BIC
fit_c_2k <- iClusterPlus(dt1 = omics_t0_2$metabo, dt2 = omics_t0_2$prot46, type = c("gaussian","gaussian"), lambda = c(0.10,0.25), K=2, maxiter = 15)
fit_c_2k$BIC
Sys.time()
```
Best with k=1 or 2 so we choose nºclusters=2
Plot heatmap:
```{r}
plotHeatmap(fit = fit_c_1k, datasets = omics_t0_2, type = c("gaussian","gaussian"))
```
It is difficult to see if clusters are according to the true label
```{r}
true_label <- as.numeric(as.factor(samples_df_T0$treatment))
icluster_cluster <- fit_c_1k$clusters
```


Now try with iClusterBayes:
```{r}
Sys.time()
#try different k values and then compare the results
#Clusters= k+1
fit_bayes_1k <- iClusterBayes(dt1 = omics_t0_2$metabo, dt2 = omics_t0_2$prot46, type = c("gaussian","gaussian"), K=1)
fit_bayes_1k$BIC
fit_bayes_2k <- iClusterBayes(dt1 = omics_t0_2$metabo, dt2 = omics_t0_2$prot46, type = c("gaussian","gaussian"), K=2)
fit_bayes_2k$BIC
Sys.time()
```

Now make plots for k=1
```{r}
plotHMBayes(fit = fit_bayes_1k, datasets = omics_t0_2,type = c("gaussian","gaussian"))
```

Plot samples in PCA space and check with correct subtypes:
```{r}
par(mar=c(5, 4, 4, 8), xpd=TRUE)
PC1 <- prcomp(omics_t0_2$metabo)

#iclusterPlus
plot(PC1$x[, 1], PC1$x[, 2], xlab = "PC 1", ylab = "PC 2", 
     main = "PC plot of samples for iClusterPlus", col= color[fit_c_1k$clusters], pch=symbols[as.numeric(as.factor(samples_df_T0$treatment))])
  
legend(x="topright", inset=c(-0.3, 0), legend= unique(samples_df_T0$treatment), title= "Treatment", col = "ivory4", pch = symbols)
legend(x="bottomright", inset=c(-0.25, 0), legend= c("Cluster 1", "Cluster 2"), title= "Clusters", col = color, pch=15)

#iclusterbayes
plot(PC1$x[, 1], PC1$x[, 2], xlab = "PC 1", ylab = "PC 2", 
     main = "PC plot of samples for iClusterBayes", col= color[fit_bayes_1k$clusters], pch=symbols[as.numeric(as.factor(samples_df_T0$treatment))])
  
legend(x="topright", inset=c(-0.3, 0), legend= unique(samples_df_T0$treatment), title= "Treatment", col = "ivory4", pch = symbols)
legend(x="bottomright", inset=c(-0.25, 0), legend= c("Cluster 1", "Cluster 2"), title= "Clusters", col = color, pch=15)
```

###10.BCC:
For this we need samples in columns so we need the data transposed
```{r}
library(bayesCC)
require(parallel)
Sys.time()
Ks <- 2:3 #try between 2 and 3 clusters to reduce computation time
names(Ks)<- paste0("K",Ks)
Results_2 <- mclapply(Ks, function(k){
  bayesCC(as.matrix(data_transposed_2_omics), K=k, IndivAlpha = T,maxiter = 10000) # we may need to add as.matrix if it does not work
})

#Compute adherence
meanAlphaStar <- c() #mean
upperAlphaStar <- c() #upper
lowerAlphaStar <- c() #lower

for (k in Ks){
  i <- names(Ks)[which(Ks==k)]
  AlphaRaw <- (Results_2[[i]]$AlphaVec[,2000:10000]-1/k) /(1-1/k)
  AlphaStar <- apply(AlphaRaw,2,mean)
  meanAlphaStar[i]<- mean(AlphaStar)
  upperAlphaStar[i]<- quantile(AlphaStar, 0.975)
  lowerAlphaStar[i]<- quantile(AlphaStar, 0.025)
}
##plot
plotCI(Ks, meanAlphaStar, ui= upperAlphaStar, li= lowerAlphaStar,
       xlab = "Number of clusters (K)", ylab = "Mean adjusted adherence")
Sys.time()
```
Almost same, so we choose 2 cluster
```{r}
K3_2 <- Results_2[[1]] #position 1
K3_2$Alpha
```
Plot the results checking source and general clustering:
```{r}
# Create function to make the points plot
plot_points <- function(PCs, cluster_col, cluster_row, color, shape) {
  for (j in 1:2) {
    points(PCs$x[K3$Cbest[, cluster_col] == 1 & K3$Lbest[[cluster_row]][, j] == 1, 1], 
           PCs$x[K3$Cbest[, cluster_col] == 1 & K3$Lbest[[cluster_row]][, j] == 1, 2], 
           col = color, pch = shape[j]) 
  }
}

colors <- c("indianred", "lightskyblue")
symbols <- c(16, 3)

plot_pcs <- function(data, title, alpha_value, m) {
  par(mar=c(5, 4, 4, 8), xpd=TRUE)
  PCs <- prcomp(t(data))
  plot(PCs$x[, 1], PCs$x[, 2], cex = 0, xlab = "PC 1", ylab = "PC 2", 
       main = bquote(.(title) ~ (alpha == .(alpha_value))))
  
  for (i in 1:2) {
    plot_points(PCs, i, m, colors[i], symbols)
  }
  legend(x="topright", inset=c(-0.3, 0), legend= unique(samples_df_T0$treatment), title= "Treatment", col = "ivory4", pch = symbols)
  legend(x="bottomright", inset=c(-0.25, 0), legend= clustersbcc, title= "Clusters", col = colors, pch=15)
}

# metabo analysis
plot_pcs(data_transposed_2_omics[[1]], "Metabo at T0", 0.653, 1)

# proteo analysis
plot_pcs(data_transposed_2_omics[[2]], "Proteo at T0", 0.884, 2)

```
##OTHER TESTS:
###CONCORDANCE (ARI):
Calculate adjusted rand index for all methods and see concordance, we want to create a matrix where both columns and rows are the different methods 
```{r}
#install.packages("cata")
library(cata) 
name_methods <- c("snf","nemo","mocluster","pinsplus","intnmf","mofa","iclusterplus","iclbayes","bayescc")
clustering_methods <- list(
  snf = group,
  nemo = clustering_nemo_T0_2,
  mocluster = cls,
  pinsplus = cluster1,
  intnmf = clusters_int,
  mofa = clusters_mofa_all$cluster,
  iclusterplus = icluster_cluster,
  iclbayes = iclusterbayes_cluster,
  bayescc = clusters_bcc
)

#Create matrix
n <- length(name_methods)
matrix_ARI <- matrix(0, nrow = n, ncol = n)
matrix_pvalue <- matrix(0, nrow = n, ncol = n)
rownames(matrix_ARI) <- name_methods
colnames(matrix_ARI) <- name_methods
rownames(matrix_pvalue) <- name_methods
colnames(matrix_pvalue) <- name_methods

#Fill matrices with ARI and pvalues
for (i in 1:n) {
  for (j in 1:n) {
    ari <- ARI(clustering_methods[[i]], clustering_methods[[j]], signif = TRUE)
    matrix_ARI[i, j] <- ari$ari
    matrix_pvalue[i,j] <- ari$pvalue
  }
}

#Print matrix
#print(matrix_ARI)
```

Now let's check the concordance between the true values:
```{r}
#install.packages("fossil")
#library("fossil")
true_clustering <- as.numeric(as.factor(samples_df_T0$treatment))

#Calculate ARI values for true clusters
matrix_ARI_true <- matrix(0, nrow = 3, ncol = n) #we want  one row for the true clustering and another for the p-value and 9 columns for the methods
rownames(matrix_ARI_true) <- c("True_cluster-ARI", "p-value","rand-index")
colnames(matrix_ARI_true) <- name_methods

#Fill matrix
for (j in 1:n){
  ari <- ARI(true_clustering, clustering_methods[[j]], signif = TRUE)
  matrix_ARI_true[1,j] <- ari$ari
  matrix_ARI_true[2,j] <- ari$pvalue
  matrix_ARI_true[3,j] <- rand.index(true_clustering, clustering_methods[[j]])
}
```

###Add metadata
- if patient is allergic or not (in mepo there are patients that are not allergic)
- comorbidities to check if that makes mepo cluster differently
All Oma has no comorbities except 3 samples, and mepo is 50-50.
```{r}
#First get common samples for the metadata
# Function to standardize names in metadata_df
standardize_names <- function(name) {
  # Use a regular expression to capture the parts and remove leading zeros
  return(sub("(O|M)-T0-0*(\\d+)", "\\1\\2", name))
}

# Apply the function to the sample names and convert the result to a vector
standard_names <- unlist(lapply(samples_df_T0$sample, standardize_names))
```

Now we have the names standarized and we want to see if there is any metadata that could be relevant to the samples not classified properly:
- load comorbidities
```{r}
comorbidities <- read.csv('Comorbilidades_Biologicos.csv',header=TRUE, sep = ';')

```

Join comorbilities to  standard names:
```{r}
samples_with_metadata <- data.frame("samples"=standard_names)
samples_with_metadata <- merge(samples_with_metadata, comorbidities, by.x = "samples", by.y = "ID", all.x = TRUE)
```

Now we can add other metadata:
- allergy, age, basal igE and Eosinophils_T0
```{r}
#Create metadata df with only desired columns
metadata_df <- data.frame("ID"=metadata$Patient_ID,"Age" = metadata$Age, "Allergic"=metadata$Allergic, "Basal_igE" = metadata$Basal_IgE, "Eosinophils"=metadata$Eosinophils_T0)
samples_with_metadata <- merge(samples_with_metadata, metadata_df, by.x = "samples", by.y = "ID", all.x = TRUE)
print(samples_with_metadata)
```

###ADDITIONAL PLOTS:
mixOmics plot for sPLS and check "clustering": we decided to plot the resulting model and see if by plotting an horizontal line we can check the division between treatments
```{r}
p <- plotIndiv(final.spls.2omics, ind.names = TRUE,  rep.space = "X-variate", group = samples_df_T0$treatment,       col.per.group = color.mixo(1:2), legend = TRUE, legend.title = 'Treatment', plot=FALSE)

p$graph + geom_hline(yintercept = 0.3, color = "red", linetype = "dashed")
```
Make correlation plots with ARI values: to have the graphically explain by adding as well the p-values
```{r}
library(corrplot)
corrplot(matrix_ARI, col = COL2('BrBG'), method = 'color', tl.col = "black", p.mat = matrix_pvalue, sig.level = -1,insig = 'p-value')
```
Plot MOFA with sample names: and check if it corresponds to the same as in mixomics, mocluster and intnmf
```{r}
par(mar=c(4, 4, 4, 8), xpd=TRUE)
color_samples <- c("darkblue","darkorange2")
plot(subset_fact[,2], subset_fact[,3], xlab="Factor 2", ylab= "Factor 3", main="Plot Factor 2 vs Factor 3 in T0", pch=16, col=colors_mofa[clusters_mofadf_all$cluster])
text(subset_fact[,2], subset_fact[,3],labels = samples_df_T0$sample,cex = 0.6, pos = 4, col = color_samples[as.numeric(as.factor(samples_df_T0$treatment))])
legend("topright", legend=c("Cluster 1","Cluster 2"), col=colors_mofa, pch=15, title="Clusters", inset=c(-0.25, 0))
```
Check common things between not-properly clustered samples:
```{r}
#Samples of mepolizumab that were not properly clustered (in moCluster, intNMF, MOFA and mixOmics)
samples_bad_df <- data.frame("sample"=c("M6","M11","M20","M30","M33","O8","O10","O12","O18","O22"))
merged_samples_bad <- merge(samples_bad_df, metadata_df,by.x = "sample", by.y = "ID", all.x = TRUE)

```


