---
title: "TCGA-BRCA"
author: "Silvia Garces"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## TCGA-BRCA DATASET:
With this dataset we intend to cluster the different samples and check if the clusters correspond to the original cancer subtypes: HER2, Basal, Luminal A and Luminal B


Load necessary libraries:
```{r}
library(ggplot2)
library(bayesCC)
library(mixOmics)
library(MOFA2)
library(NEMO)
library(SNFtool)
library(omicade4)
library(mogsa)
library(IntNMF)
library(PINSPlus)
library(iClusterPlus)

```

The data has the samples as columns and the variables as rows. 
First we need to read the data (from bayesCC package):
```{r BRCAData}
data(BRCAData)
#[mrna, dna methylation, mirna, protein]
```

This data contains 348 samples.
We load the clustering done by TCGA so we can compare our results:
```{r}
table_TCGA_subtypes <- read_delim("TFM_PRACTICAS/table_TCGA_subtypes.csv", delim = ";", escape_double = FALSE, trim_ws = TRUE)
subtype <- table_TCGA_subtypes$Cluster #get subtype
ids <- table_TCGA_subtypes$ID
ids <- gsub("-", ".", as.character(ids))
namesExp <- colnames(BRCAData[[1]])
namesExp2 <- substr(namesExp,1,12)
MatchSub <- match(namesExp2, ids, nomatch = 0) #match names
sum(ids[MatchSub] == namesExp2)
subtype <- subtype[MatchSub] #get original groups (1, 2, 3, 4)
subtypes <- c("HER2","Basal","Luminal A", "Luminal B") #get labels for each group
```


##Start methods
###1. SNF: 
For this method we need to separate the data into 4 and the calculate the distance for each of the matrices
Main problem --> We have to transpose the matrices as SNF needs the samples to be in rows 
```{r}
Sys.time() #to calculate running time

Dist1 = (dist2(as.matrix(t(BRCAData[[1]])), as.matrix(t(BRCAData[[1]]))))^(1/2)
Dist2 = (dist2(as.matrix(t(BRCAData[[2]])), as.matrix(t(BRCAData[[2]]))))^(1/2)
Dist3 = (dist2(as.matrix(t(BRCAData[[3]])), as.matrix(t(BRCAData[[3]]))))^(1/2)
Dist4 = (dist2(as.matrix(t(BRCAData[[4]])), as.matrix(t(BRCAData[[4]]))))^(1/2)

K=20 #number of neighbors
sigma=0.5

W1 = affinityMatrix(Dist1, K, sigma)
W2 = affinityMatrix(Dist2, K, sigma)
W3 = affinityMatrix(Dist3, K, sigma)
W4 = affinityMatrix(Dist4, K, sigma) 


W_joined <- SNF(list(W1,W2,W3,W4), K, 10) #generate model

estimateNumberOfClustersGivenGraph(W_joined, 2:5)
Sys.time()
```
Select number of optimal clusters = 2
```{r}
C=2 #number of clusters
group=spectralClustering(W_joined,C) #generate clusters
displayClusters(W_joined,group)
```
Check with true labels and plot:
```{r}
M_label=cbind(group, subtype)
colnames(M_label)=c("spectralClustering","TrueLabel")
rownames(M_label)=namesExp
M_label_colors=cbind("spectralClustering"=getColorsForGroups(M_label[,"spectralClustering"],
colors=c("blue","red")),"TrueLabel"=getColorsForGroups(M_label[,"TrueLabel"], colors = c("green","violet","orange","grey3")))

#Make plot
layout(matrix(c(1,2), ncol=2, byrow=TRUE))
par(mar=c(5, 4, 4, 2))
displayClustersWithHeatmap(W_joined, group, M_label_colors)
par(mar=c(0,0,0,0))
plot(1, type = "n", axes=F, xlab="", ylab="") # Create empty plot
legend(x="top", legend=unique(group), fill=unique(M_label_colors[,"spectralClustering"]), title = "SNF clusters", bg="transparent", horiz=TRUE)
legend(x="center", legend=subtype_labels, fill= c("green","violet","orange","grey3"), title = "TCGA subtypes", bg="transparent", horiz=TRUE)

```

Compare with TCGA subtypes:
```{r}
# Assign labels
subtype_labels <- c("Her2", "Basal", "LumA", "LumB")
cluster_labels_snf <- c("Cluster 1", "Cluster 2")

Clusters_snf <- rep(1, 348)
Clusters_snf[group == 2] <- 2
#Clusters_nemo[cls == 3] <- 3
#Clusters_nemo[cls == 4] <- 4
# ConfusionMat: All
confmat_snf <- matrix(nrow = 4, ncol = 2)
for (i in 1:4) {
  for (j in 1:2) {
    confmat_snf[i, j] <- sum(subtype == i & Clusters_snf == j)
  }
}
# Assign names
rownames(confmat_snf) <- subtype_labels
colnames(confmat_snf) <- cluster_labels_snf
```

###2. NEMO:
With this method we will compute the affinity graph, we can use data without transposing - samples in columns
```{r}
Sys.time() #calculate running time
affinity_graph <- nemo.affinity.graph(BRCAData, k= 20) #k corresponds with number of neighbors to make matrix
num_clusters <- nemo.num.clusters(affinity_graph)
clustering_nemo <- nemo.clustering(BRCAData, num.clusters = num_clusters, num.neighbors = 5)
Sys.time()


```
Plot heatmap with different colors for clusters
```{r}
NEMO_label=cbind(clustering_nemo, subtype)
colnames(NEMO_label)=c("nemoClustering","TrueLabel")
rownames(NEMO_label)=namesExp
NEMO_label_colors=cbind("nemoClustering"=getColorsForGroups(NEMO_label[,"nemoClustering"],
colors=c("blue","red")),"TrueLabel"=getColorsForGroups(NEMO_label[,"TrueLabel"], colors = c("green","violet","orange","grey3")))


displayClustersWithHeatmap(affinity_graph, clustering_nemo, NEMO_label_colors)
plot.new()
legend(x="top", legend=unique(group), fill=unique(NEMO_label_colors[,"nemoClustering"]), title = "NEMO clusters", bg="transparent", horiz=TRUE, cex=0.8)
legend(x="center", legend=subtype_labels, fill= c("green","violet","orange","grey3"), title = "TCGA subtypes", bg="transparent", horiz=TRUE, cex=0.8)

```

Compare clusterings with TCGA subtypes
```{r}
# Assign labels
subtype_labels <- c("Her2", "Basal", "LumA", "LumB")
cluster_labels_nemo <- c("Cluster 1", "Cluster 2")

Clusters_nemo <- rep(1, 348)
Clusters_nemo[clustering_nemo == 2] <- 2
#Clusters_nemo[cls == 3] <- 3
#Clusters_nemo[cls == 4] <- 4
# ConfusionMat: All
confmat_nemo <- matrix(nrow = 4, ncol = 2)
for (i in 1:4) {
  for (j in 1:2) {
    confmat_nemo[i, j] <- sum(subtype == i & Clusters_nemo == j)
  }
}
# Assign names
rownames(confmat_nemo) <- subtype_labels
colnames(confmat_nemo) <- cluster_labels_nemo
```

###3. MCIA:
For this method the data can be used as it is as the samples are already in columns
```{r}
Sys.time() #calculate computation time
mcoin <- mcia(BRCAData)
Sys.time()

```

Now make the plots: these ones are difficult to understand
```{r}
plot.mcia(mcoin,axes = 1:2 , phenovec = subtype, sample.lab = FALSE, df.color = 1:4)
plotVar(mcoin, var=NA, bg.var.col=1:4, var.lab=TRUE)
```
Make plots using 2 latent variables to check possible clusters
```{r}
color <- c("olivedrab","burlywood","deepskyblue","maroon")
subset_mcia <- mcoin$mcoa$SynVar
par(mar=c(5, 4, 4, 8), xpd=TRUE)
plot(subset_mcia[, 1], subset_mcia[,2], xlab= "Variable 1", ylab= "Variable 2", col=color[subtype], pch=20, main = "Plot of samples")
legend("right", legend=subtype_labels, col=color,pch=20, title="TCGA subtypes", inset=c(-0.3, 0))
```
###4. moCluster:
No need to modify the data as samples need to be in columns
```{r}
Sys.time() #for the running time
#Make model using 10 comps
moa <- mbpca(BRCAData, ncomp=10, k="all", method = "globalScore", option = "lambda1",center=TRUE, scale= FALSE, moa= TRUE, svd.solver = "fast", maxiter = 1000)
plot(moa, value="eig", type=2)

#Search for appropriate number of components
r <- bootMbpca(moa, mc.cores = 1, B=20, replace = FALSE, resample = "sample")
```
Choose best number of components = 3
```{r}
#Create model for 3 components
moa_3var <- mbpca(BRCAData, ncomp=3, k=0.1, method = "globalScore", option = "lambda1",center=TRUE, scale= FALSE, moa= TRUE, svd.solver = "fast", maxiter = 1000)

scr <- moaScore(moa)
scrs <-moaScore(moa_3var)

#do correlation
diag(cor(scr[, 1:3], scrs))

#Do hierarchical cluster and plot
hcl <- hclust(dist(scrs))
cls <- cutree(hcl, k=4) #Select number of clusters, we want 4
Sys.time()

#Plot the results obtained with 3 components
clsColor <- as.factor(cls)
levels(clsColor) <- c("red","blue","orange", "pink")
clsColor<- as.character((clsColor))
```
Plot heatmap and add legend:
```{r}
heatmap(t(scrs[hcl$order, ]), ColSideColors = clsColor[hcl$order], Rowv = NA, Colv=NA)
legend(x= "right", legend = c("Cluster 1","Cluster 2", "Cluster 3", "Cluster 4"), col = c("red","blue","orange", "pink"), pch= 15)
```
Plot samples in PCA space with subtypes and cluster:
```{r}
symbols <- c(16, 3, 8, 5)
par(mar=c(5, 4, 4, 8), xpd=TRUE)
plot(scrs[, 1:2], col=clsColor, pch=symbols[subtype], main = "PC1 and PC2")
legend("topright", legend=c("Cluster 1","Cluster 2", "Cluster 3", "Cluster 4"), col=unique(clsColor), pch=15, title="Clusters", inset=c(-0.25, 0))
legend("bottomright", legend=subtype_labels, col=c("lightslategrey"), pch=symbols, title="Subtypes", inset=c(-0.25, 0))

par(mar=c(5, 4, 4, 8), xpd=TRUE)
plot(scrs[, 2:3], col=clsColor, pch=symbols[subtype], main = "PC2 and PC3")
legend("topright", legend=c("Cluster 1","Cluster 2", "Cluster 3", "Cluster 4"), col=unique(clsColor), pch=15, title="Clusters", inset=c(-0.25, 0))
legend("bottomright", legend=subtype_labels, col=c("lightslategrey"), pch=symbols, title="Subtypes", inset=c(-0.25, 0))

```

Check clusters with TCGA subtypes:
```{r}
# Assign labels
subtype_labels <- c("Her2", "Basal", "LumA", "LumB")
cluster_labels <- c("Cluster 1", "Cluster 2", "Cluster 3", "Cluster 4")

Clusters_mo <- rep(1, 348)
Clusters_mo[cls == 2] <- 2
Clusters_mo[cls == 3] <- 3
Clusters_mo[cls == 4] <- 4
# ConfusionMat: All
confmat_mo <- matrix(nrow = 4, ncol = 4)
for (i in 1:4) {
  for (j in 1:4) {
    confmat_mo[i, j] <- sum(subtype == i & Clusters_mo == j)
  }
}
# Assign names
rownames(confmat_mo) <- subtype_labels
colnames(confmat_mo) <- cluster_labels
```

###5. PINSPlus:
We need to modify the data so samples are in rows- transpose data
```{r}
Sys.time()
data1<- t(BRCAData[[1]])
data2<- t(BRCAData[[2]])
data3<- t(BRCAData[[3]])
data4<- t(BRCAData[[4]])
dataList <- list(data1,data2,data3,data4) #make list with transposed data
#Use following function as we have multi-omics data
pins_cluster <- SubtypingOmicsData(dataList =dataList , clusteringMethod = "kmeans", kMax=4) #Select number of clusters

Sys.time()

#Two  clustering outputs
cluster1 <- pins_cluster$cluster1
cluster2 <- pins_cluster$cluster2

# Plot the clusters for Stage I using a scatter plot
plot(cluster1, pch = 16, col = "blue", main = "Cluster Plot (Stage I)", xlab = "Sample", ylab = "Cluster")
# Plot the clusters for Stage II using a scatter plot
plot(cluster2, pch = 16, col = "red", main = "Cluster Plot (Stage II)", xlab = "Sample", ylab = "Cluster")
```
Plot PCA for first 2 components to check distribution of samples:
```{r}
# Subset the first two columns of the 'x' matrix (which are the PCA)
subset_x <- pins_cluster[["dataTypeResult"]][[1]][["pca"]][["x"]]
cluster_colors <- c("coral", "cadetblue3") #There are only two clusters
symbols <- c(16, 3, 8, 5)
# Plot PC1 vs PC2
par(mar=c(5, 4, 4, 8), xpd=TRUE)
plot(subset_x[, 1], subset_x[, 2], 
     xlab="PC1", ylab="PC2", 
     main="Plot of samples in PC1 vs PC2", 
     pch=symbols[subtype], col=cluster_colors[cluster1])
legend("topright", legend=c("Cluster 1","Cluster 2"), col=cluster_colors, pch=15, title="Clusters", inset=c(-0.25, 0))
legend("bottomright", legend=subtype_labels, col="lightslategrey", pch=symbols, title="Subtypes", inset=c(-0.25, 0))
```

Compare cluster obtained with TCGA subtypes:
```{r}
# Assign labels
subtype_labels <- c("Her2", "Basal", "LumA", "LumB")
cluster_labels <- c("Cluster 1", "Cluster 2", "Cluster 3", "Cluster 4")

Clusters_pins <- rep(1, 348)
Clusters_pins[cluster1 == 2] <- 2
Clusters_pins[cluster1 == 3] <- 3
Clusters_pins[cluster1 == 4] <- 4
# ConfusionMat: All
confmat_pins <- matrix(nrow = 4, ncol = 4)
for (i in 1:4) {
  for (j in 1:4) {
    confmat_pins[i, j] <- sum(subtype == i & Clusters_pins == j)
  }
}
# Assign names
rownames(confmat_pins) <- subtype_labels
colnames(confmat_pins) <- cluster_labels

```


###6. intNMF:
Samples should be on rows so we need to modify the data, moreover we need to make the data positive, as the program does not work with negative values.Use data from PINSPlus (already transposed)

```{r}
Sys.time() #running time
#check if there is any negative value
if (!all(data1>=0)){
  data1_pos <- pmax(data1 + abs(min(data1)), .Machine$double.eps)
} else {
  data1_pos <- data1
}
if (!all(data2>=0)){
  data2_pos <- pmax(data2 + abs(min(data2)), .Machine$double.eps)
} else {
  data2_pos <- data2
}
if (!all(data3>=0)){
  data3_pos <- pmax(data3 + abs(min(data3)), .Machine$double.eps)
} else {
  data3_pos <- data3
}
if (!all(data4>=0)){
  data4_pos <- pmax(data4 + abs(min(data4)), .Machine$double.eps)
} else {
  data4_pos <- data4
}

dat <- list(data1_pos,data2_pos,data3_pos,data4_pos) #make list with positive data

#Get optimum number of clusters: nmf.opt.k
opt_k<- nmf.opt.k(dat=dat, n.runs = 5, n.fold = 5, k.range = 2:7, result = TRUE, 
                  make.plot = TRUE, progress = TRUE)

Sys.time()
```
Check number of optimal clusters=4
```{r}
#We check and the optimal is 4
fit <- nmf.mnnals(dat=dat, k=4, maxiter = 200, st.count = 20, n.ini = 15, ini.nndsvd = TRUE, seed = TRUE)

#Plot matrix
ConsensusMatPlot(fit, colLab = TRUE, rowLab = TRUE)
```


Plot the clusters obtained in factorial space:
```{r}
clusters_int <- fit$clusters
subsets_df <- fit$W #get the basis matrix for the different datasets
clus_colors <- c("darkorange1", "deepskyblue", "gold", "lightpink3")
symbols <- c(16, 3, 8, 5)
# Plot Factor 1 vs Factor 2
par(mar=c(5, 4, 4, 8), xpd=TRUE)
plot(subsets_df[, 1], subsets_df[, 2], 
     xlab="Factor 1", ylab="Factor 2", 
     main="Plot of Factor 1 vs Factor 2", 
     pch=symbols[subtype], col=clus_colors[clusters_int])
legend("topright", legend=c("Cluster 1","Cluster 2", "Cluster 3", "Cluster 4"), col=clus_colors, pch=15, title="Clusters", inset=c(-0.25, 0))
legend("bottomright", legend=subtype_labels, col="lightslategrey", pch=symbols, title="Subtypes", inset=c(-0.25, 0))

# Plot Factor 3 vs Factor 4
par(mar=c(5, 4, 4, 8), xpd=TRUE)
plot(subsets_df[, 3], subsets_df[, 4], 
     xlab="Factor 3", ylab="Factor 4", 
     main="Plot of Factor 3 vs Factor 4", 
     pch=symbols[subtype], col=clus_colors[clusters_int])
legend("topright", legend=c("Cluster 1","Cluster 2", "Cluster 3", "Cluster 4"), col=clus_colors, pch=15, title="Clusters", inset=c(-0.25, 0))
legend("bottomright", legend=subtype_labels, col="lightslategrey", pch=symbols, title="Subtypes", inset=c(-0.25, 0))
```

Compare plots with TCGA subtype:
```{r}
# Assign labels
subtype_labels <- c("Her2", "Basal", "LumA", "LumB")
cluster_labels <- c("Cluster 1", "Cluster 2", "Cluster 3", "Cluster 4")

Clusters_nmf <- rep(1, 348)
Clusters_nmf[fit$cluster == 2] <- 2
Clusters_nmf[fit$cluster == 3] <- 3
Clusters_nmf[fit$cluster == 4] <- 4
# ConfusionMat: All
confmat_nmf <- matrix(nrow = 4, ncol = 4)
for (i in 1:4) {
  for (j in 1:4) {
    confmat_nmf[i, j] <- sum(subtype == i & Clusters_nmf == j)
  }
}
# Assign names
rownames(confmat_nmf) <- subtype_labels
colnames(confmat_nmf) <- cluster_labels
```

###7. MixOmics:
For this we need to specify the names of the different omics, moreover samples are on rows so we need to transpose the data
```{r}
X1 <- t(BRCAData[[1]]) #mrna
X2 <- t(BRCAData[[2]]) #dna meth
X3 <- t(BRCAData[[3]]) #mirna
X4 <- t(BRCAData[[4]]) #protein
X<- list(mrna=X1, dnamet=X2, mirna=X3, protein=X4)
 
```

For these data we need to use the unsupervised method, with multi-block PLS using all the different variables
```{r}
Sys.time() #for the running time
block.pls.result <- block.pls(X,indY=4, design="full")
Sys.time()
plotIndiv(block.pls.result, group = subtype_labels[subtype], ind.names= FALSE, legend = TRUE, legend.title = "Subtypes")
plotVar(block.pls.result, legend = TRUE)

```

Now we select the variables by using Multi-block sPLS:
```{r}
##Perform multiblock sPLS: we set the number of features to use
#If we use indY we do not use keepY, only keepX so we need to make sure to include all the variables in the list
Sys.time() #for the running time
list.keepX <- list(mrna=rep(10,2), dnamet=rep(10,2),mirna=rep(10,2), protein=rep(10,2))
#list.keepY <- c(rep(10,2))

block.spls.result <- block.spls(X,indY=4, design="full",keepX = list.keepX)
Sys.time()

###plot
plotIndiv(block.spls.result,group = subtype_labels[subtype], ind.names= FALSE, legend = TRUE, legend.title = "Subtypes")
plotVar(block.spls.result, title= "PLS Correlations - no cutoff",legend = TRUE)
plotVar(block.spls.result, title= "PLS Correlations - 0.5 cutoff",cutoff=0.5, legend = TRUE)
```
Try doing supervised learning as we know the true values, for this we need to have a train and test dataset so we would need to separate the data into two. 
```{r}
Sys.time() #running time
Y_label <- subtype_labels[subtype]
design = matrix(0.1, ncol = length(X), nrow = length(X), 
                dimnames = list(names(X), names(X)))
diag(design) = 0 # set diagonal to 0s
basic.diablo.model = block.splsda(X = X, Y = Y_label, ncomp = 5, design = design)
# run component number tuning with repeated CV
perf.diablo = perf(basic.diablo.model, validation = 'Mfold', 
                   folds = 10, nrepeat = 10)
plot(perf.diablo)

#Set optimal number of components
ncomp <- perf.diablo$choice.ncomp$WeightedVote["Overall.BER", "centroids.dist"]
# set grid of values for each component
test.keepX = list (mrna = rep(10,2), 
                   dnamet = rep(10,2),
                   mirna = rep(10,2),
                   protein= rep(10,2))

# run the feature selection tuning - takes a long time
#tune.TCGA = tune.block.splsda(X = X, Y = Y_label, ncomp = ncomp, 
#                              test.keepX = test.keepX, design = design,
#                              validation = 'Mfold', folds = 10, nrepeat = 1,
#                              dist = "centroids.dist")
#list.keepX = tune.TCGA$choice.keepX # set the optimal values of features to retain
#Create final model
final.diablo.model = block.splsda(X = X, Y = Y_label, ncomp = ncomp, 
                          keepX = test.keepX, design = design)
Sys.time()
```
Make plots for DIABLO's output
```{r}
plotDiablo(final.diablo.model, ncomp = 1)
plotIndiv(final.diablo.model, ind.names = FALSE, legend = TRUE, 
          title = 'Plot for each dataset')
plotArrow(final.diablo.model, ind.names = FALSE, legend = TRUE, 
          title = 'DIABLO')
plotLoadings(final.diablo.model, comp = 1, contrib = 'max', method = 'median') # to check contribution of features for each dataset
```
Heatmap: this is done in the console
X11()
cimDiablo(final.diablo.model, comp = 1, legend.position = "right")

Now check performance and AUC:
```{r}
# run repeated CV performance evaluation
perf.diablo = perf(final.diablo.model, validation = 'Mfold', 
                   M = 10, nrepeat = 10, 
                   dist = 'centroids.dist') 

perf.diablo$MajorityVote.error.rate
```

Plot AUC for each dataset:
```{r}
auc.splsda = auroc(final.diablo.model, roc.block = "mrna", 
                   roc.comp = 2, print = FALSE)
auc.splsda = auroc(final.diablo.model, roc.block = "dnamet", 
                   roc.comp = 2, print = FALSE)
auc.splsda = auroc(final.diablo.model, roc.block = "mirna", 
                   roc.comp = 2, print = FALSE)
auc.splsda = auroc(final.diablo.model, roc.block = "protein", 
                   roc.comp = 2, print = FALSE)
```

###8. MOFA
We do not need to modify the structure of the data
```{r}
Sys.time()
data_mofa = BRCAData
names(data_mofa) <- c("mrna", "dnamet", "mirna", "prot")
mofa_obj <- create_mofa(data_mofa)
plot_data_overview(mofa_obj)
#Specify the options
##Default
data_opts <- get_default_data_options(mofa_obj)

model_opts <- get_default_model_options(mofa_obj)
model_opts$num_factors <- 10

train_opts <- get_default_training_options(mofa_obj)
train_opts$seed <- 42

#Start training the model
mofa_obj <- prepare_mofa(mofa_obj, data_options = data_opts,
                            model_options = model_opts, training_options = train_opts)
outfile = file.path(getwd(),"model_15_4.hdf5") #save model for future use
MOFA_object_trained <- run_mofa(mofa_obj, outfile, use_basilisk=TRUE)
Sys.time()
```

Generate plots and check factors:
```{r}
plot_factor_cor(MOFA_object_trained)
##max_r2= maximum percentage of variance
plot_variance_explained(MOFA_object_trained, max_r2 = 25)
###Factor 1 is the one with more explained variance for all the different datasets

##total variance
plot_variance_explained(MOFA_object_trained, plot_total = T)[[2]]
```

Plot the weights to see which variables have higher weights in the factor
```{r}
plot_weights(MOFA_object_trained, view = "mrna", factor= 1, nfeatures = 10)
plot_weights(MOFA_object_trained, view = "dnamet", factor= 1, nfeatures = 10)
plot_weights(MOFA_object_trained, view = "mirna", factor= 1, nfeatures = 10)
plot_weights(MOFA_object_trained, view = "prot", factor= 1, nfeatures = 10)
```
Plot heatmaps
```{r}
plot_data_heatmap(MOFA_object_trained, view = "mrna", factor = 1, features = 25,
                  denoise = TRUE, cluster_rows= TRUE, cluster_cols= TRUE, 
                  show_rownames= TRUE, show_colnames=FALSE  ,scale="row")
plot_data_heatmap(MOFA_object_trained, view = "dnamet", factor = 1, features = 25,
                  denoise = TRUE, cluster_rows= TRUE, cluster_cols= TRUE, 
                  show_rownames= TRUE, show_colnames=FALSE  ,scale="row")
plot_data_heatmap(MOFA_object_trained, view = "mirna", factor = 1, features = 25,
                  denoise = TRUE, cluster_rows= TRUE, cluster_cols= TRUE, 
                  show_rownames= TRUE, show_colnames=FALSE  ,scale="row")
plot_data_heatmap(MOFA_object_trained, view = "prot", factor = 1, features = 25,
                  denoise = TRUE, cluster_rows= TRUE, cluster_cols= TRUE, 
                  show_rownames= TRUE, show_colnames=FALSE  ,scale="row")
```
Generate clusters and plot them:
```{r}
set.seed(32)
clusters_mofa_all <- cluster_samples(MOFA_object_trained, k=4, factor="all") #needs to select clusters beforehand
clusters_mofadf_all <- data.frame(cluster=clusters_mofa_all$cluster)
ggplot(clusters_mofadf_all, aes(x = factor(cluster), fill = factor(cluster))) + 
  geom_bar() +
  labs(title = "Cluster Distribution for all factors",
       x = "Cluster",
       y = "Count") +
  scale_fill_discrete(name = "Cluster")

clusters_mofa_1 <- cluster_samples(MOFA_object_trained, k=4, factor=1)
clusters_mofadf_1 <- data.frame(cluster=clusters_mofa_1$cluster)
ggplot(clusters_mofadf_1, aes(x = factor(cluster), fill = factor(cluster))) + 
  geom_bar() +
  labs(title = "Cluster Distribution for Factor 1",
       x = "Cluster",
       y = "Count") +
  scale_fill_discrete(name = "Cluster")


```
Plot factor 1 vs factor 3:
```{r}
subset_fact <- MOFA_object_trained@expectations$Z[["group1"]]
colors_mofa <- c("darkorchid1","darkseagreen4","lightsalmon1", "powderblue")
par(mar=c(5, 4, 4, 8), xpd=TRUE)
plot(subset_fact[,1], subset_fact[,3], xlab="Factor1", ylab= "Factor 3", main="Plot Factor 1 vs Factor 3",
     pch=symbols[subtype], col=colors_mofa[clusters_mofadf_1$cluster])
legend("topright", legend=c("Cluster 1","Cluster 2", "Cluster 3", "Cluster 4"), col=colors_mofa, pch=15, title="Clusters", inset=c(-0.25, 0))
legend("bottomright", legend=subtype_labels, col="lightslategrey", pch=symbols, title="Subtypes", inset=c(-0.25, 0))
```
Now with ggplot
```{r}
ggplot(data = as.data.frame(subset_fact), aes(x = Factor1, y = Factor3)) +
  geom_point(aes(shape=as.factor(subtype), color=as.factor(clusters_mofadf_1$cluster))) +
  scale_shape_manual(values=symbols, labels = subtype_labels)+
  labs(x = "Factor1", y = "Factor3", title = "Plot Factor 1 vs Factor 3", color = "Clusters", shape = "Subtypes")
```

Compare clusters with TCGA subtypes
```{r}
# Assign labels
subtype_labels <- c("Her2", "Basal", "LumA", "LumB")
cluster_labels <- c("Cluster 1", "Cluster 2", "Cluster 3", "Cluster 4")

Clusters_mofa <- rep(1, 348)
Clusters_mofa[clusters_mofadf_1$cluster == 2] <- 2
Clusters_mofa[clusters_mofadf_1$cluster == 3] <- 3
Clusters_mofa[clusters_mofadf_1$cluster == 4] <- 4
# ConfusionMat: All
confmat_mofa <- matrix(nrow = 4, ncol = 4)
for (i in 1:4) {
  for (j in 1:4) {
    confmat_mofa[i, j] <- sum(subtype == i & Clusters_mofa == j)
  }
}
# Assign names
rownames(confmat_mofa) <- subtype_labels
colnames(confmat_mofa) <- cluster_labels

```

###9. BCC: 
There is no need to modify the data as it is originally from this package
We found that the best are 4 clusters so we need to modify the code.
This takes approximately 1 hour to run
```{r}
require(parallel)
Sys.time()
Ks <- 2:4
names(Ks)<- paste0("K",Ks)
Results <- mclapply(Ks, function(k){
  bayesCC(BRCAData, K=k, IndivAlpha = T,maxiter = 10000)
})

#Compute adherence
meanAlphaStar <- c() #mean
upperAlphaStar <- c() #upper
lowerAlphaStar <- c() #lower

for (k in Ks){
  i <- names(Ks)[which(Ks==k)]
  AlphaRaw <- (Results[[i]]$AlphaVec[,2000:10000]-1/k) /(1-1/k)
  AlphaStar <- apply(AlphaRaw,2,mean)
  meanAlphaStar[i]<- mean(AlphaStar)
  upperAlphaStar[i]<- quantile(AlphaStar, 0.975)
  lowerAlphaStar[i]<- quantile(AlphaStar, 0.025)
}
##plot
plotCI(Ks, meanAlphaStar, ui= upperAlphaStar, li= lowerAlphaStar,
       xlab = "Number of clusters (K)", ylab = "Mean adjusted adherence")
Sys.time()
```
Best with 4 clusters
```{r}
K3 <- Results[[3]] #position 3 of the results array
K3$Alpha
```

Plot the results checking source and general clustering:
```{r}
# Create function to make the points plot
plot_points <- function(PCs, cluster_col, cluster_row, color, shape) {
  for (j in 1:4) {
    points(PCs$x[K3$Cbest[, cluster_col] == 1 & K3$Lbest[[cluster_row]][, j] == 1, 1], 
           PCs$x[K3$Cbest[, cluster_col] == 1 & K3$Lbest[[cluster_row]][, j] == 1, 2], 
           col = color, pch = shape[j]) 
  }
}

colors <- c("black", "indianred", "lightskyblue", "mediumseagreen")
symbols <- c(16, 3, 8, 5)

plot_pcs <- function(data, title, alpha_value, m) {
  par(mar=c(5, 4, 4, 8), xpd=TRUE)
  PCs <- prcomp(t(data))
  plot(PCs$x[, 1], PCs$x[, 2], cex = 0, xlab = "PC 1", ylab = "PC 2", 
       main = bquote(.(title) ~ (alpha == .(alpha_value))))
  
  for (i in 1:4) {
    plot_points(PCs, i, m, colors[i], symbols)
  }
  legend(x="topright", inset=c(-0.3, 0), legend= subtypes, title= "TCGA Subtypes", col = "ivory4", pch = symbols)
  legend(x="bottomright", inset=c(-0.25, 0), legend= clustersbcc, title= "Clusters", col = colors, pch=15)
}

# GE analysis
plot_pcs(BRCAData[[1]], "GE", 0.86, 1)

# DNA methylation analysis
plot_pcs(BRCAData[[2]], "ME", 0.57, 2)

# miRNA analysis
plot_pcs(BRCAData[[3]], "miRNA", 0.65, 3)

# Protein analysis
plot_pcs(BRCAData[[4]], "RPPA", 0.67, 4)
```
Check original data with our clusters
```{r}
Clusters <- rep(1, 348)
Clusters[K3$Cbest[, 2] == 1] <- 2
Clusters[K3$Cbest[, 3] == 1] <- 3
Clusters[K3$Cbest[, 4] == 1] <- 4
# ConfusionMat: All
confmat <- matrix(nrow = 4, ncol = 4)
for (i in 1:4) {
  for (j in 1:4) {
    confmat[i, j] <- sum(subtype == i & Clusters == j)
  }
}
rownames(confmat) <- c("Her2","Basal","LumA","LumB")
colnames(confmat) <- c("Cluster 1", "Cluster 2","Cluster 3","Cluster 4")
```


###10. iClusterPlus:
We need to modify the data as samples need to be in rows, we can use X1,X2,X3 and X4 from mixOmics as we already transposed that data. First we need to determine the distribution of each dataset. However for this method to work we need to center the data, to do so we need to scale it
```{r}
sc_X2 <- scale(X2)
sc_X3 <- scale(X3)

hist(X1) #gaussian
hist(sc_X2) #gaussian
hist(sc_X3) #gaussian
hist(X4) #gaussian
```

Once we get the data type (assume Gaussian)

```{r}
Sys.time()
#try different k values and then compare the results
fit_c_2 <- iClusterPlus(dt1 = X1, dt2 = sc_X2, dt3 = sc_X3, dt4 = X4, type = c("gaussian","gaussian","gaussian","gaussian"), lambda = c(0.90,0.60,0.75,0.05), K=2, maxiter = 10)
fit_c_2$BIC
fit_c_3 <- iClusterPlus(dt1 = X1, dt2 = sc_X2, dt3 = sc_X3, dt4 = X4, type = c("gaussian","gaussian","gaussian","gaussian"), lambda = c(0.90,0.60,0.75,0.05), K=3, maxiter = 10)
fit_c_3$BIC
Sys.time()
```
All BIC values are the same so we choose 4 clusters to check with the subtypes. Then we will plot the heatmap.
We need to load library gplots to use it
```{r}
plotHeatmap(fit = fit_c_3, datasets = list(X1,sc_X2,sc_X3,X4), type = c("gaussian","gaussian","gaussian","gaussian"))
```

Now lets try with iClusterBayes
```{r}
Sys.time()
fit_bayes_2 <- iClusterBayes(dt1 = X1, dt2 = sc_X2, dt3 = sc_X3, dt4 = X4, type = c("gaussian","gaussian","gaussian","gaussian"), K=2)
fit_bayes_2$BIC
fit_bayes_3 <- iClusterBayes(dt1 = X1, dt2 = sc_X2, dt3 = sc_X3, dt4 = X4, type = c("gaussian","gaussian","gaussian","gaussian"), K=3)
fit_bayes_3$BIC
Sys.time()
```
Plot heatmaps, the lowest BIC is obtained with k=3, so we plot the heatmap:
```{r}
plotHMBayes(fit = fit_bayes_3, datasets = list(X1,sc_X2,sc_X3,X4),type = c("gaussian","gaussian","gaussian","gaussian"))
```
Plot samples in PCA space as done in BCC and check with correct subtypes:
```{r}
par(mar=c(5, 4, 4, 8), xpd=TRUE)
PC1 <- prcomp(X1)
PC2 <- prcomp(sc_X2)
PC3 <- prcomp(sc_X3)
PC4 <- prcomp(X4)
#iclusterPlus
plot(PC1$x[, 1], PC1$x[, 2], xlab = "PC 1", ylab = "PC 2", 
     main = "PC plot of samples for iClusterPlus", col= color[fit_c_3$clusters], pch=symbols[subtype])
  
legend(x="topright", inset=c(-0.3, 0), legend= subtypes, title= "TCGA Subtypes", col = "ivory4", pch = symbols)
legend(x="bottomright", inset=c(-0.25, 0), legend= clustersbcc, title= "Clusters", col = color, pch=15)

#iclusterbayes
plot(PC1$x[, 1], PC1$x[, 2], xlab = "PC 1", ylab = "PC 2", 
     main = "PC plot of samples for iClusterBayes", col= color[fit_bayes_3$clusters], pch=symbols[subtype])
  
legend(x="topright", inset=c(-0.3, 0), legend= subtypes, title= "TCGA Subtypes", col = "ivory4", pch = symbols)
legend(x="bottomright", inset=c(-0.25, 0), legend= clustersbcc, title= "Clusters", col = color, pch=15)
```


Check the clusters and compare them with original TCGA 
```{r}
# Assign labels
subtype_labels <- c("Her2", "Basal", "LumA", "LumB")
cluster_labels <- c("Cluster 1", "Cluster 2", "Cluster 3", "Cluster 4")

##IclusterPlus
Clusters_icluster <- rep(1, 348)
Clusters_icluster[fit_c_3$clusters == 2] <- 2
Clusters_icluster[fit_c_3$clusters == 3] <- 3
Clusters_icluster[fit_c_3$clusters == 4] <- 4
# Confusion matrix
confmat_icluster <- matrix(nrow = 4, ncol = 4)
for (i in 1:4) {
  for (j in 1:4) {
    confmat_icluster[i, j] <- sum(subtype == i & Clusters_icluster == j)
  }
}
# Assign names
rownames(confmat_icluster) <- subtype_labels
colnames(confmat_icluster) <- cluster_labels

##IclusterBayes
Clusters_iclusterB <- rep(1, 348)
Clusters_iclusterB[fit_bayes_3$clusters == 2] <- 2
Clusters_iclusterB[fit_bayes_3$clusters == 3] <- 3
Clusters_iclusterB[fit_bayes_3$clusters == 4] <- 4
# Confusion matrix
confmat_iclusterB <- matrix(nrow = 4, ncol = 4)
for (i in 1:4) {
  for (j in 1:4) {
    confmat_iclusterB[i, j] <- sum(subtype == i & Clusters_iclusterB == j)
  }
}
# Assign names
rownames(confmat_iclusterB) <- subtype_labels
colnames(confmat_iclusterB) <- cluster_labels

```

##Calculate rand index:
To check concordance between the methods we will calculate the rand index, to do so we will only check the methods that gave as a clustering output
As we have different number of clusters for some of the methods the resulting matrix won't be symmetrical.
First try all methods
```{r}
#install.packages("cata")
library(cata) 
icluster_cluster <- fit_c_3$clusters
iclusterbayes_cluster <- fit_bayes_3$clusters
name_methods <- c("snf","nemo","mocluster","pinsplus","intnmf","mofa","iclusterplus","iclbayes","bayescc")
clustering_methods <- list(
  snf = group,
  nemo = clustering_nemo,
  mocluster = unname(cls),
  pinsplus = cluster1,
  intnmf = clusters_int,
  mofa = clusters_mofadf_1$cluster,
  iclusterplus = icluster_cluster,
  iclbayes = iclusterbayes_cluster,
  bayescc = clusters_bcc
)

#Create matrix
n <- length(name_methods)
matrix_ARI <- matrix(0, nrow = n, ncol = n)
matrix_pvalue <- matrix(0, nrow = n, ncol = n)
rownames(matrix_ARI) <- name_methods
colnames(matrix_ARI) <- name_methods
rownames(matrix_pvalue) <- name_methods
colnames(matrix_pvalue) <- name_methods

#Fill matrices with ARI and pvalues
for (i in 1:n) {
  for (j in 1:n) {
    set.seed(333)
    ari <- ARI(clustering_methods[[i]], clustering_methods[[j]], signif = TRUE)
    matrix_ARI[i, j] <- ari$ari
    matrix_pvalue[i,j] <- ari$pvalue
  }
}
```

Now try only the ones that generated 2 clusters
```{r}
cluster_2_methods <- list(
  snf = group,
  nemo = clustering_nemo,
  pinsplus = cluster1
)
#Create matrix
n <- length(cluster_2_methods)
matrix_ARI_2_clusters <- matrix(0, nrow = n, ncol = n)
matrix_pvalue_2_clusters <- matrix(0, nrow = n, ncol = n)
rownames(matrix_ARI_2_clusters) <- names(cluster_2_methods)
colnames(matrix_ARI_2_clusters) <- names(cluster_2_methods)
rownames(matrix_pvalue_2_clusters) <- names(cluster_2_methods)
colnames(matrix_pvalue_2_clusters) <- names(cluster_2_methods)

#Fill matrices with ARI and pvalues
for (i in 1:n) {
  for (j in 1:n) {
    ari <- ARI(cluster_2_methods[[i]], cluster_2_methods[[j]], signif = TRUE)
    matrix_ARI_2_clusters[i, j] <- ari$ari
    matrix_pvalue_2_clusters[i,j] <- ari$pvalue
  }
}
```

Now for all the ones with 4 clusters:
```{r}
clustering_methods_4 <- list(
  mocluster = unname(cls),
  intnmf = clusters_int,
  mofa = clusters_mofadf_1$cluster,
  iclusterplus = icluster_cluster,
  iclbayes = iclusterbayes_cluster,
  bayescc = clusters_bcc
)

#Create matrix
n <- length(clustering_methods_4)
matrix_ARI_4_clusters <- matrix(0, nrow = n, ncol = n)
matrix_pvalue_4_clusters <- matrix(0, nrow = n, ncol = n)
rownames(matrix_ARI_4_clusters) <- names(clustering_methods_4)
colnames(matrix_ARI_4_clusters) <- names(clustering_methods_4)
rownames(matrix_pvalue_4_clusters) <- names(clustering_methods_4)
colnames(matrix_pvalue_4_clusters) <- names(clustering_methods_4)

#Fill matrices with ARI and pvalues
for (i in 1:n) {
  for (j in 1:n) {
    ari <- ARI(clustering_methods_4[[i]], clustering_methods_4[[j]], signif = TRUE)
    matrix_ARI_4_clusters[i, j] <- ari$ari
    matrix_pvalue_4_clusters[i,j] <- ari$pvalue
  }
}
```

Now calculate the ARI and rand index value considering the true clustering
```{r}
#install.packages("fossil")
#library("fossil")
#subtype variable are the true clusters
n <- length(name_methods)
#Calculate ARI values for true clusters
matrix_ARI_true <- matrix(0, nrow = 3, ncol = n) #we want one row for the ARI value, another for the p-value and other for the rand.index and 9 columns for the methods
rownames(matrix_ARI_true) <- c("True_cluster-ARI", "p-value","rand-index")
colnames(matrix_ARI_true) <- name_methods

#Fill matrix
for (j in 1:n){
  ari <- ARI(subtype, clustering_methods[[j]], signif = TRUE)
  matrix_ARI_true[1,j] <- ari$ari
  matrix_ARI_true[2,j] <- ari$pvalue
  matrix_ARI_true[3,j] <- rand.index(subtype, clustering_methods[[j]])
}
```

Now I want to calculate the adjusted mutual information (AMI):
```{r}
#install.packages("aricode")
#library(aricode)
matrix_AMI <-  matrix(0, nrow = 1, ncol = n) #ami between true clustering and methods clustering
rownames(matrix_AMI) <- "AMI- true cluster"
colnames(matrix_AMI) <- name_methods

#Fill matrix
for (j in 1:n){
  ami <- AMI(subtype, clustering_methods[[j]])
  matrix_AMI[1,j] <- ami
}
```

Now do the same for all methods:
```{r}
matrix_AMI_full <-  matrix(0, nrow = n, ncol = n) #ami between methods
rownames(matrix_AMI_full) <- name_methods
colnames(matrix_AMI_full) <- name_methods

#Fill matrix
for (i in 1:n){
  for (j in 1:n){
  ami <- AMI(clustering_methods[[i]], clustering_methods[[j]])
  matrix_AMI_full[i,j] <- ami
  }
}

```

###Additional plots:
ARI in correlation matrix:
```{r}
corrplot(matrix_ARI, col = COL2('BrBG'), method = 'color', tl.col = "black", p.mat = matrix_pvalue, sig.level = -1,insig = 'p-value')
corrplot(matrix_ARI_2_clusters, col = COL2('BrBG'), method = 'color', tl.col = "black", p.mat = matrix_pvalue_2_clusters, sig.level = -1,insig = 'p-value')
corrplot(matrix_ARI_4_clusters, col = COL2('BrBG'), method = 'color', tl.col = "black", p.mat = matrix_pvalue_4_clusters, sig.level = -1,insig = 'p-value')
```
AMI in correlation plot:
```{r}
#pdf("prueba.pdf")
corrplot(matrix_AMI_full, addCoef.col = 'black',col = COL2('BrBG'), method = 'color', tl.col = "black")
#dev.off()
```


